---
title: "List_of_Commands"
author: "Laurine Cabiling"
date: "2023-01-23"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Intro
This is a list of commands from lecture, excluding commands we may have used independtly from homework assignments. Not all the codes are runable because there is no connected files.

# Lab 1 commands
```{r}
#get working directory
#getwd()
#set working directory 
#setwd()
```

```{r}
#making a vector or string of values into an object 
# c means concatenate
x <- c(3,43,7,8,534)
```

```{r}
#mean of object
mean(x)
```

```{r}
#median of object
median(x)
```

## "#" = titles 
## = reduced font size, the more # = reduced size
smallest font size 
_italics_
**bold**
line break = two places 

## Links 
[website name]
(website link)
(mailto: email address)

```{r}
#input code chunk with mac is ...
#`option+command+i`
```

```{r}
#install.packages("name of package")
#library("name of package") to download data every session
```

```{r}
#ggplot(data, subset(x = factor(cyl))) + geom_bar()
#to create a plot graph?
```

#Lab 2 commands

```{r}
#assigning values to an object 
x <- 34
s <- 44
```

```{r}
#printing object 
x
```

```{r}
#suming objects 
distance <- sum(x,s)
```

## Types of Data
```{r}
#Five frequently used classes of data: 
  # 1. numeric
  # 2. integer - adding an L automatically denotes an integer
  # 3. character
  # 4. logical
  # 5. complex
my_numeric <- 43
my_integer <- 2L 
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```

```{r}
#find class of function
class(my_numeric)
```

```{r}
# is my numeric an integer? 
is.integer(my_numeric)
```

```{r}
# create a new object specified as a different class
my_integer <- as.integer(my_numeric)
```

```{r}
# is there NA in my data - says which data is T/F
is.na(x)
```

```{r}
# are there any NAs in my data - T/F
anyNA(x)
```

```{r}
#to remove any NAs when calculating
# na.rm = T
mean(x, na.rm = T)
```

data structures include...
  vectors
  lists
  matrices
  data frames
  factors
  
```{r}
#numeric vector
my_vector <- c(20,34,23)
```

```{r}
#character vector
days_of_the_week <- c("sunday", "monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday")
```

```{r}
# generate a sequence of numbers
my_vector_sequence <- c(1:100)
```

```{r}
# use [] to pull out elements in a vector
days_of_the_week[3]
```

```{r}
# pull out data greater than ...
my_vector_sequence > 80

# pull out data greater than or equal to ...
my_vector_sequence >= 49

# pull out data less than ...
my_vector_sequence < 30

# pull out data less than or equal to...
my_vector_sequence <= 43

# pull out data equal to ...
my_vector_sequence == 39

# Use [] to get the values, not the logical evaluation
my_vector_sequence[my_vector_sequence <= 20]
```

## How to do Data Matrices 

```{r}
Philosophers_Stone <- c(317.5, 657.1)
Chamber_of_Secrets <- c(261.9, 616.9)
Prisoner_of_Azkaban <- c(249.5, 547.1)
Goblet_of_Fire <- c(290.0, 606.8)
Order_of_the_Phoenix <- c(292.0, 647.8)
Half_Blood_Prince <- c(301.9, 632.4)
Deathly_Hallows_1 <- c(295.9, 664.3)
Deathly_Hallows_2 <- c(381.0, 960.5)
```

```{r}
#combine all objects into one
box_office <- c(Philosophers_Stone, Chamber_of_Secrets, Prisoner_of_Azkaban, Goblet_of_Fire, Order_of_the_Phoenix, Half_Blood_Prince, Deathly_Hallows_1, Deathly_Hallows_2)
```

```{r}
#create the matrix, and organize using `nrow` and `byrow`
harry_potter_matrix <- matrix(box_office, nrow = 8, byrow = T)
```

```{r}
#name the vectors for the rows and columns 
region <- c("US", "non-US")
titles <- c("Philosophers_Stone", "Chamber_of_Secrets", "Prisoner_of_Azkaban", "Goblet_of_Fire", "Order_of_the_Phoenix", "Half_Blood_Prince", "Deathly_Hallows_1", "Deathly_Hallows_2")
```

```{r}
#name the columns
colnames(harry_potter_matrix) <- region
#name the rows
rownames(harry_potter_matrix) <- titles
```

```{r}
# print the matrix
harry_potter_matrix
```

```{r}
#sum the rows 
global <- rowSums(harry_potter_matrix)
```

```{r}
#add new column to show calculation
all_harry_potter_matrix <- cbind(harry_potter_matrix, global)
```

```{r}
#sum the columns
final_earnings <- colSums(all_harry_potter_matrix)
```

```{r}
# pull out data, the first value is the column, then the row
harry_potter_matrix[2,1]
# pull out data, `:` selects specified elements in a column
harry_potter_matrix[1:4]
```

```{r}
#calculate an entire row or column
non_us_earnings <- all_harry_potter_matrix[ ,2]
mean(non_us_earnings)
```

# Lab 3 commands 
Building data frames

```{r}
Sex <- c("Male", "Female", "Male")
Length <- c(3.2, 3.7, 3.4)
Weight <- c(2.9, 4.0, 3.1)
```

```{r}
#get the data frame
hbirds <- data.frame(Sex, Length, Weight)
```

```{r}
# column names
names(hbirds)
```

```{r}
# dimensions of data frame
dim(hbirds)
```

```{r}
# structure of data frame
str(hbirds)
```

```{r}
# switching to lowercase
hbirds <- data.frame(sex = Sex, length_in = Length, weight_oz = Weight)
```

```{r}
#Select values in an entire column using `$` sign. 
w <- hbirds$weight_oz
```

```{r}
#add new data using `rbind()`, bind new vector to data frame row-wise
new_bird <- c("Female", 3.6, 3.9)
hbirds <- rbind(hbirds, new_bird)
```

```{r}
# adding columns
hbirds$neighborhood <- c("Lakewood", "Brentwood", "Lakewood", "Scenic Heights")
```

```{r}
#Saving our dataframe as a csv
write.csv(hbirds, "hbirds_data.csv", row.names = FALSE)
```

```{r}
#Opening a .csv file & give it a name
hbirds <- readr::read_csv("hbirds_data.csv")
```

```{r}
#Looking at the data frame in different ways

#structure
#str()

#a type of summary
#glimpse()

#a type of summary
#summary()

#number of rows or observations
#nrow()

#number of columns or variables 
#ncol()

#dimensions
#dim()

#names of column
#name()

#prints the first n rows of the data frame
#head(data, n = )

#prints the last n rows of the data frame
#tail(data, n = )

#table - useful when limited number of categorical variables
#table(data$variable)

#See the data frame in the separate tab
#View()
```

```{r}
#Looking at a class within the data frame
class(hbirds$neighborhood)
#looking at the level within the data frame
levels(hbirds$neightborhood)
#changing the class of data
hbirds$neighborhood <- as.factor(hbirds$neighborhood)
```

```{r}
#subsets are to pull out observations that meet a critera in a variable
# little_fish <- subset(fish, length <= 110)
```

##Lab 4 commands

```{r}
#select is to filter but with columns instead of rows
#select(fish, "lakeid", "scalelength")
```

```{r}
#to add range of columns
#select(fish, fish_id:length)
```

```{r}
# use - operator to select everything except these variables
#select(fish, -fish_id, -annumber, -length)
```

```{r}
#select columns that contain a character string
#select(fish, contains("length"))

#select columns that start with this name
#select(fish, starts_with("radii"))

#select columns that end with a charater string
#select(fish, ends_with())

#select columns that match a regular expression
#select(fish, matches())

#select column names that are from a group of names
#select(fish, one_of())

#a column contains a letter (in this case "a") followed by a subsequent string (in this case "er").  
#select(fish,matches("a.+er"))

#select column based on class of data
#select_if(fish, is.numeric)

#select all columns that are NOT a class of data, add `~`
#select_if(fish,~is.numeric(.))
```

```{r}
#renaming columns
#mammals_new <- rename(mammals, genus = "Genus",wean_mass = "wean mass", max_life = "max. life", litter_size = "litter size", litters_per_year = "litters/year")

#to lower case all of the columns
#mammals <- select_all(mammals, tolower)

#to upper case all of the columns
#mammals <- select_all(mammals, toupper)

#to remove the blank spaces
#select_all(mammals, ~str_replace(., " ","_"))
#select_all(mammals, ~str_replace(., "[ ._/]", "" ))
```

```{r}
# filter extracts data that meets specific criteria within a variable
#filter(fish, lakeid == "AL")

# filter allows to use the expected operators to filter through data ;
  # >, >=, <, <=, != (not equal), and == (equal)

# filter multiple values within same variable
#filter(fish, length %in% c(167,175))

#between
#filter(fish, between(scalelength, 2.5, 3))

#extract observations "near" a certain value, and specify a tolerance
#filter(fish, near(radii_length_mm, 2, tol = 0.2))

#filter to extract data based on multiple conditions
#filter(fish, lakeid == "AL", length == 350)

# `|` operator means or 
#filter(fish, lakeid == "AL" | lenght > 350)

#will return rows where both conditions are met
#filter(condition1, condition2)

# will return all rows where condition one is true but condition 2 is not.  
#filter(condition1, !condition2)

# will return rows where condition 1 and/or condition 2 is met.  
#filter(condition1 | condition2)
  
# will return all rows where only one of the conditions is met, and not when both conditions are met. 
#filter(xor(condition1, condition2))
```






