---
title: "Lab 9 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Seating
1. If you already have a group, it will be helpful to start sitting next to each other.
2. If you don't have a group, hang tight we are working on it.
3. Please setup your computer as normal for today.

## Warm-up
1. In the data folder there is an epidemiology data set on an outbreak of malaria.
```{r}
epidem <- readr::read_csv("data/malaria.csv")
epidem
```

2. `rdt` refers to rapid diagnostic test and they are identified here by age group.

3. Make the data tidy and store them as a new object.
```{r}
library(tidyverse)
library(janitor)
```
```{r}
new_epidem <- janitor::clean_names(epidem)
new_epidem
```
pivot longer 
```{r}
names(new_epidem)
```
the rdt are data = need to make it tidy --> `pivot_longer` 
```{r}
new_epidem_long <- new_epidem %>% 
  pivot_longer(cols = starts_with("malaria_rdt"),
               names_to = "age_class",
               values_to = "cases") %>% 
  select(newid, data_date, submitted_date, location_name, province, district, age_class, cases)
```

4. Which district had the highest *total* number of cases on July 30, 2020?
```{r}
new_epidem_long %>% 
  filter(data_date == "2020-07-30") %>%
  group_by(district) %>% 
  summarise(total_cases = sum(cases, na.rm = T)) %>% 
  arrange(-total_cases)
```
