---
title: "Lab 7 Homework"
author: "Laurine Cabiling"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
``` 

```{r}
#install.packages("naniar")
library("naniar")
```

## Data
**1. For this homework, we will use two different data sets. Please load `amniota` and `amphibio`.**  

`amniota` data:  
Myhrvold N, Baldridge E, Chan B, Sivam D, Freeman DL, Ernest SKM (2015). “An amniote life-history
database to perform comparative analyses with birds, mammals, and reptiles.” _Ecology_, *96*, 3109.
doi: 10.1890/15-0846.1 (URL: https://doi.org/10.1890/15-0846.1).
```{r}
amniota <- readr::read_csv("data/amniota.csv")
```

`amphibio` data:  
Oliveira BF, São-Pedro VA, Santos-Barrera G, Penone C, Costa GC (2017). “AmphiBIO, a global database
for amphibian ecological traits.” _Scientific Data_, *4*, 170123. doi: 10.1038/sdata.2017.123 (URL:
https://doi.org/10.1038/sdata.2017.123).
```{r}
amphibio <- readr::read_csv("data/amphibio.csv")
```

## Questions  
**2. Do some exploratory analysis of the `amniota` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.**  
```{r}
glimpse(amniota)
```

```{r}
hist(amniota$fledging_age_d)
```

**3. Do some exploratory analysis of the `amphibio` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.** 
The NA's are represented as NA's. 
```{r}
glimpse(amphibio)
```
```{r}
anyNA(amphibio)
```

**4. How many total NA's are in each data set? Do these values make sense? Are NA's represented by values?**   

`amniota`  
The NA's are represented as -999 values in this dataframe. 
```{r}
amniota %>% 
 summarise_all(~(sum(is.na(.))))
```
```{r}
amniota %>% 
  count("-999")
```

`amphibio`  
The NA's appear as NA's in the dataframe 
```{r}
amphibio %>% 
  summarise_all(~(sum(is.na(.))))
```

**5. Make any necessary replacements in the data such that all NA's appear as "NA".**   
```{r}
amniota <- amniota %>% 
  na_if("-999")
```

**6. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amniota` data.**  
```{r}
naniar::miss_var_summary(amniota)
```

**7. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amphibio` data.**
```{r}
naniar::miss_var_summary(amphibio)
```

**8. For the `amniota` data, calculate the number of NAs in the `egg_mass_g` column sorted by taxonomic class; i.e. how many NA's are present in the `egg_mass_g` column in birds, mammals, and reptiles? Does this results make sense biologically? How do these results affect your interpretation of NA's?**  
```{r}
names(amniota)
```
```{r}
amniota %>% 
  group_by(class) %>% 
  summarize(total_NA = sum(is.na(egg_mass_g)))
```
```{r}
amniota %>% 
 summarise_all(~(sum(is.na(.))))
```
```{r}
amniota %>% 
  group_by(class) %>% 
  summarize(total=n())
```
To ensure I got the right number of individual species were missing an egg_mass_g value, I summarized all of the missing NA's. Then to see if the results made biological sense, I summarized the total number of individuals within each class. The values make sense almost all of Mammalia do not lay eggs, so it makes sense the entire class is missing an egg weight value. Furthermore, we know reptiles and birds do lay eggs, but not all of the species were able to get an egg weight for some reason. Therefore, we can be missing important data as the mass of the egg can indicate possible future health risks for the off springs as it can be an indication that there is a decrease in nutrients within their habitat. 

**9. The `amphibio` data have variables that classify species as fossorial (burrowing), terrestrial, aquatic, or arboreal.Calculate the number of NA's in each of these variables. Do you think that the authors intend us to think that there are NA's in these columns or could they represent something else? Explain.**
```{r}
names(amphibio)
```
```{r}
amphibio %>% 
  select(Fos, Ter, Aqu, Arb) %>% 
  summarize_all((~(sum(is.na(.)))))
```
```{r}
amphibio %>% 
  select(Species, Fos, Ter, Aqu, Arb)
```
I think the authors did mean to put NA's in the values to indicate that the species was never documented or found in that specific habitat. It could be like a binary code where 1 is yes, they are found there, and NA instad of 0 to mean no they were not found there. 

**10. Now that we know how NA's are represented in the `amniota` data, how would you load the data such that the values which represent NA's are automatically converted?**
```{r}
amniota <- readr::read_csv("data/amniota.csv") %>% 
  na_if("-999")
```
## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  