---
title: "Lab 6 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Setup
1. Login to the lab computer (please don't use your personal computer).  
2. Navigate to github.com and login.   
2. Use GitHub desktop to clone your repository to the desktop.   
5. Copy the class repository to the desktop (https://github.com/jmledford3115/datascibiol).  
6. Copy the files for today's lab from the class repository and paste them into **your** repository.  
7. Open today's lab in RStudio.  

## Review from last time
### *With a partner, discuss the following questions*
1. What is a pipe? Why are they useful in R?
A pipe is this ` %>% ` symbol and it pulls the output from one function into the input of another function.Makes it faster to combine data. 

2. What is the shortcut for making a pipe?
Shift + Cmnd + M

3. What is the difference between `select`, `filter`, and `mutate`?
select is for variables, filter is for observations, and mutate is to change a part of the object to another one. 

4. What are the logistics of midterm 1?

### Warm-up
1. Open the data `ecs21351-sup-0003-SupplementS1.csv`
```{r}
library("tidyverse")
library("janitor")
```
```{r}
ecosphere <- read.csv("data/ecs21351-sup-0003-SupplementS1.csv", skip =2)
```

2. Clean the names of the variables
```{r}
names(ecosphere)
```

```{r}
ecosphere <- ecosphere %>%  clean_names()
```

3. Explore the data using a function of your choice
```{r}
glimpse(ecosphere)
```

4. Are there more veggie, omni, insecivore, or nectar feeders in the data?
```{r}
ecosphere %>% 
  tabyl (diet) %>% 
  arrange(desc(n))
```
There are more insectivore feeders among the list of feeders in the data. 

5. Which families are Veggie?
```{r}
ecosphere %>% 
  filter(diet == "Vegetation") %>% 
  select(family, diet) %>% 
  tabyl(diet)
```

